--- 
title: "README" 
output: github_document 
--- 

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE) 
``` 
## Introduction 
Modelling disease risk and survival using longitudinal risk factor trajectories
is of interest in various clinical scenarios. We propose a dynamic risk score modeling
framework for multiple longitudinal risk factors and survival in the presence
of dependent censoring, where both events depend on participants’ post-baseline clinical progression. Our model requires relatively few random effects regardless of the number of longitudinal risk factors and can therefore accommodate multiple longitudinal risk factors in a parsimonious manner.

## Installation 
The development version of package `DRS` can be installed using: 
```{r eval=FALSE} 
library(devtools) 
install_github("chzhang25/DRS", quiet = TRUE) 
``` 

## Example 
```{r eval=FALSE} 
library(DRS) 
## load required data
data(simda)
head(da.long)
da.id <- subset(da.long, v.time == 0)

## fit the proposed model
out <- DRS.JM(coxForm = ~ Lt1 + Lt2 + Lt3 + Lt4 + Lt5,
              jmFixedForm = ~ v.time + W,
              jmRandomForm = ~ v.time | id,
              jmCoxForm1 = Surv(obsT, delta == 1) ~ W,
              jmCoxForm2 = Surv(obsT, delta == 2) ~ W,
              timeVar = "v.time", data.long = da.long, data.id = da.id,
              coxControl = list(n.iter = 30))
out
# Call:
# DRS.JM(coxForm = ~Lt1 + Lt2 + Lt3 + Lt4 + Lt5, jmFixedForm = ~v.time + 
#     W, jmRandomForm = ~v.time | id, jmCoxForm1 = Surv(obsT, delta == 
#     1) ~ W, jmCoxForm2 = Surv(obsT, delta == 2) ~ W, timeVar = "v.time", 
#     data.long = da.long, data.id = da.id, coxControl = list(n.iter = 30))
# 
#                     Estimate StdErr  z value Pr(>|z|)
# betas.(Intercept)_1   0.2249 0.0890   2.5259   0.0115
# betas.v.time_1        0.1948 0.0237   8.2344   0.0000
# betas.W_1            -0.0892 0.0498  -1.7905   0.0734
# betas.(Intercept)_2  -0.2327 0.0899  -2.5895   0.0096
# betas.v.time_2        0.4136 0.0304  13.6221   0.0000
# betas.W_2             0.0587 0.0487   1.2059   0.2279
# log.sigma1           -1.2379 0.0904 -13.6898   0.0000
# log.sigma2           -1.1423 0.1149  -9.9391   0.0000
# gammas1.W            -0.2387 0.0760  -3.1428   0.0017
# alpha1                1.4017 0.1055  13.2812   0.0000
# gammas2.W             0.0640 0.0765   0.8365   0.4029
# alpha2                1.1514 0.1183   9.7292   0.0000
# Lt1_1                 0.3160 0.0551   5.7304   0.0000
# Lt2_1                 0.5173 0.0288  17.9356   0.0000
# Lt3_1                -0.4094 0.0531  -7.7037   0.0000
# Lt4_1                 0.4774 0.0209  22.8330   0.0000
# Lt5_1                 0.4868 0.0303  16.0643   0.0000
# Lt1_2                 0.3101 0.0747   4.1529   0.0000
# Lt2_2                 0.4945 0.0360  13.7499   0.0000
# Lt3_2                -0.4230 0.0606  -6.9789   0.0000
# Lt4_2                 0.2544 0.0328   7.7450   0.0000
# Lt5_2                 0.6447 0.0339  18.9984   0.0000
``` 
`betas.(Intercept)_1`, `betas.v.time_1` and `betas.W_1` are the estimates for fixed effects $\boldsymbol{\beta_1}$ of the longitudinal risk score of Event 1, whereas `betas.(Intercept)_2`, `betas.v.time_2` and `betas.W_2` are the estimates for fixed effects $\boldsymbol{\beta_2}$ of longitudinal risk score of Event 2.

`log.sigma1` and `log.sigma2` are the estimates of $log(\sigma_1)$ and $log(\sigma_2)$ separately.

`gammas1.W` and `gammas2.W` are the estimates for $\boldsymbol{\gamma_1}$ and $\boldsymbol{\gamma_2}$ separately. `alpha1` and `alpha2` are the estimates for $\boldsymbol{\alpha_1}$ and $\boldsymbol{\alpha_2}$ separately.

`Lt1_1`, `Lt2_1`, `Lt3_1`, `Lt4_1` and `Lt5_1` are the estimates of the components in $\boldsymbol{\xi_1}$, whereas `Lt1_2`, `Lt2_2`, `Lt3_2`, `Lt4_2` and `Lt5_2` are the estimates of the components in $\boldsymbol{\xi_2}$.

`gammas.bs.11 - gammas.bs.19` (not printed by default) are estimates for approximating log baseline hazards $h_{01}(t)$, whereas `gammas.bs.21 - gammas.bs.29` (not shown) are estimates for approximating log baseline hazards $h_{02}(t)$.

`D1- D10` (not printed by default) are estimates for Choleski factors of the random effects covariance matrix ($\textbf{D}$).

## Reference 
Cuihong Zhang, Jing Ning, Jianwen Cai, James E. Squires, Steven H. Belle & Ruosha Li (2023). Dynamic Risk Score Modeling for Multiple Longitudinal Risk Factors and Survival [Manuscript submitted for publication]
